---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #3 - instructions

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

# Lab 3: Avenger's Peril

## As a team

Extract from the data below two data sets in long form `deaths` and `returns`

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```

Get the data into a format where the five columns for Death[1-5] are replaced by two columns: Time, and Death. Time should be a number between 1 and 5 (look into the function `parse_number`); Death is a categorical variables with values "yes", "no" and "". Call the resulting data set `deaths`. 

Similarly, deal with the returns of characters.

Based on these datasets calculate the average number of deaths an Avenger suffers.

## Data Tidying

```{r}
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv", stringsAsFactors = FALSE)
head(av)
```


```{r}
library(dplyr)
library(tidyr)
library(readr)

deaths <- av %>%
  rowwise() %>%  
  mutate(
    Time = sum(c_across(starts_with("Death")) == "YES", na.rm = TRUE),
    #if they died more than they returned, then the Death column is "YES"
    Returned = if_else( Time > sum(c_across(starts_with("Return")) == "YES", na.rm = TRUE), "NO", "YES"),
    Death = if_else(Returned == "YES", "NO", "YES"))%>%
    #end Death calc  
  ungroup() %>%
  select(-Death1, -Death2, -Death3, -Death4, -Death5, -Return1, -Return2, -Return3, -Return4, -Return5)

deaths
```

Average deaths suffered for an Avenger:
```{r}
mean(deaths$Time)
```

## Individually

For each team member, copy this part of the report. 

Each team member picks one of the statements in the FiveThirtyEight [analysis](https://fivethirtyeight.com/features/avengers-death-comics-age-of-ultron/) and fact checks it based on the data. Use dplyr functionality whenever possible.

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.


## Team Member Hudson Nebbe


### FiveThirtyEight Statement

> Out of 173 listed Avengers, my analysis found that 69 had died at least one time after they joined the team.5 That’s about 40 percent of all people who have ever signed on to the team.

### Include the code

``` {r}
hnn <- deaths %>% filter(Time > 0) %>% count()
#number of Time rows greater than 0
hnn

#number of Time rows greater than 0 divided by number of avengers
hnn/173

```

### Include your answer

Using the dplyr filter and count methods, I was able to determine that the number 69 does make sense according to the article. After that I checked to make sure the percentage that the author gave made sense. This statement seems factually backed!

## Team Member Huy Nguyen

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Team Member Jack Olsan

### FiveThirtyEight Statement

> Quote the statement you are planning to fact-check.

### Include the code

Make sure to include the code to derive the (numeric) fact for the statement

### Include your answer

Include at least one sentence discussing the result of your fact-checking endeavor.

Upload your changes to the repository. Discuss and refine answers as a team.

## Team Member Madhu Avula

### FiveThirtyEight Statement

> “The MVP of the Earth-616 Marvel Universe Avengers has to be Jocasta — an android based on Janet van Dyne and built by Ultron — who has been destroyed five times and then recovered five times.”

### Include the code

```{r}
jocasta_data <- av %>%
  filter(grepl("Jocasta", Name.Alias))  # Adjust if necessary for exact match

# Count Jocasta’s deaths
deaths_count <- jocasta_data %>%
  select(starts_with("Death")) %>%
  summarise(total_deaths = sum(. == "YES", na.rm = TRUE))

# Count Jocasta’s returns
returns_count <- jocasta_data %>%
  select(starts_with("Return")) %>%
  summarise(total_returns = sum(. == "YES", na.rm = TRUE))

deaths_count$total_deaths
returns_count$total_returns
```

### Include your answer
Based on the analysis, Jocasta has indeed been recorded as having 5 total deaths and 5 returns in the Avengers data set. This confirms the FiveThirtyEight statement about Jocasta being destroyed five times and recovered five times. The statement is accurate based on the data set provided.

Upload your changes to the repository. Discuss and refine answers as a team.
